---
title: "Descriptive Analysis of Grade Outcomes"
subtitle: "Proposal Final Project"
author: "The Null Wranglers Team"
format: html
editor: visual
---

```{r load_pkgs, message=FALSE, output=FALSE}
#| code-fold: true
#| code-summary: "Installed Packages"
# Load Libraries
library(knitr)
if (!require(pacman))
  install.packages(pacman)

pacman::p_load(tidyverse,   # Data wrangling
               dlookr,      # Exploratory Data Analysis
               formattable, # Present neat table format
               gt,          # Alternating formatting for the tables
               gtsummary)   # Summary for the tables
```

## Introduction

Low student success rates have a profound impact on an educational institution's academic offerings, their ability to provide financial support and scholarships, as well as their overall professional reputation. Much is unknown about how the corona-virus pandemic has impacted student success in higher education. Thus, in order to better understand this impact, this study seeks to conduct a detailed examination of the grade value outcomes during this time. By examining the grade value outcomes for the years before, during and after the pandemic, patterns and trends may be discovered that might help guide student support services in the future. The analysis will breakdown the grade values outcomes by college, department, division level and course level using clustering and classification techniques. Anomalies will be isolated for a more fine grained inspection that includes course attributes. These course attributes will be used in an effort to create a model for predictive grade value outcomes.

## Goals

The high-level goal of this study is to analyze and understand the grade outcomes at the University of Arizona (UofA) pre-, during, and after the COVID-19 pandemic. This investigation revolves around two primary questions:

1.  Pattern Analysis -- What if any patterns or relationships exist between colleges, departments, or course level and grade outcomes for enrolled students?

2.  Predictive Modeling -- Can a predictive model be developed that accurately forecasts grade outcomes based on course attributes?

## Dataset

The first data set used for this analysis, is known as the DEW Rates, contains information about the breakdown of grade value outcomes during the academic years of 2018-19, 2019-20, and 2020-21 at the University of Arizona. The consists of 35697 observations of courses with 23 variables related to enrollment and grade values. Below is a description of the data:

| Variable                     | Description                                                                                                                                                          |
|------------------------------|------------------------------------------|
| College                      | Identifies the college at the University of Arizona                                                                                                                  |
| Department                   | Identifies which department within given college                                                                                                                     |
| Subject.Code                 | Identifies the subject of the course with a 3 or 4 letter code                                                                                                       |
| Catalog.Number               | Identifies the catalog number of the course                                                                                                                          |
| Course.Description           | Identifies the title or the course                                                                                                                                   |
| Course.Level                 | Identifies if the course is lower, upper or graduate division                                                                                                        |
| Total.Student.Count          | The total enrollment for the course                                                                                                                                  |
| D_GRADE_COUNT                | The number of enrolled students that earned a D grade in the course                                                                                                  |
| FAIL_GRADE_COUNT             | The number of enrolled students that earned a E(Fail) grade in the course                                                                                            |
| WITHDRAW_GRADE_COUNT         | The number of enrolled students that withdrew from the course                                                                                                        |
| DEW_COUNT                    | The total of D, E, and W grades earned by enrolled students in the course                                                                                            |
| PASS_GRADE_COUNT             | The number of enrolled students that received a C or greater grade value                                                                                             |
| WITHDRAW_FULLMED_GRADE_COUNT | The number of enrolled students that took a full medical withdrawal                                                                                                  |
| INCOMPLETE_UNGRADED_COUNT    | The number of enrolled students that did not receive grades                                                                                                          |
| TERM_LD                      | The semester that the course was offered                                                                                                                             |
| ACAD_YR_SID                  | The academic year that the course was offered                                                                                                                        |
| Percent.D.Grade              | The percentage of enrolled students that received a D grade in the course                                                                                            |
| Percent.E.Grade              | The percentage of enrolled students that received a E grade in the course                                                                                            |
| Percent.W.Grade              | The percentage of enrolled students that received a W grade in the course                                                                                            |
| Percent.DEW                  | The percentage of enrolled students that received a D, E, or W grade in the courseThe percentage of enrolled students that received a D, E, or W grade in the course |
| Percent.Passed               | The percentage of enrolled students that received a C or greater grade value in the course                                                                           |
| Per.Full..Medical.Withdrawal | The percentage of enrolled students that had a full medical withdraw in the course                                                                                   |
| Per.Ungraded..Incomplete     | The percentage of enrolled students that did not receive a grade value in the course                                                                                 |

: Grade Outcomes Data

```{r  message=FALSE}
#| label: load-dataset
# read in data
dew_counts <- read.csv("data/DEW Rates.csv")

kable(head(dew_counts))
```

The second data set used for this analysis, is known as the SY20 and SY21, contains information about course attributes for the academic years of 2019-20, and 2020-21 at the University of Arizona. Unfortunately the academic year of 2018-19 was unable to obtained for this study. There are over 260,000 observations of courses with 27 variable per observation.

| Variable          | Description                                                       |
|-----------------------|-------------------------------------------------|
| Term              | The term the course was offered                                   |
| Campus            | The campus the course was offered at                              |
| Session           | The type of session for the course - regular, first 7 weeks, etc, |
| Subject           | The 3 or 4 letter subject code of the course                      |
| Cat . .           | The course catalog code                                           |
| Section           | The section number of the course                                  |
| Class . .         | The class identifier code                                         |
| Start.Date        | The start date of the course                                      |
| End.Date          | The end of the course                                             |
| Meet . .          | The number of meetings for each course                            |
| Req.Desig         | A course type identifier                                          |
| P.F. Opt          | A flag that indicates if the course was a P or F option           |
| Component         | Category/Type of course                                           |
| Units             | The number of units of the course                                 |
| Min.Units         | The minimum number of units allowed for the course                |
| Max.Units         | The maximum number of units allowed for the course                |
| Course            | The name of the course                                            |
| Combined.Section  | Identifier if the course is combined with another course          |
| Meeting.Days      | The list of days the courses meets on                             |
| Start             | The start time of day that the course meets                       |
| End               | The end time of day that the course meets                         |
| Facility          | The location or room of the course meetings                       |
| Total.Enroll      | The total number of students enrolled                             |
| Max.Enroll        | The maximum number of students allowed to enrolled                |
| Rm.Cap            | The maximum number of students allowed in the meeting space       |
| Enrollment.Status | Identifier if the course was an open or closed enrollment         |
| Mode              | The modelity of the course                                        |

: Course Schedule Overview

```{r message=FALSE}
#| label: load-dataset1
# read in two csv files
sy20_data <- read.csv("data/Course Schedule Overview - SY20.csv")

kable(head(sy20_data))
```

## Analysis plan

***Data Collection***

-    Historical Grade Data: Collect historical grade data from the University of Arizona for the years before, during, and after the COVID-19 pandemic. This data should include information on grade value outcomes for courses by semester.

-   Course Attributes Data: Gather data on course attributes such as class size, instructional format (online, hybrid, in-person), faculty information, time and day of the class offering, and any other relevant features that might influence grade outcomes.

***Exploratory Data Analysis (EDA)***

2.1 Descriptive Statistics

Calculate summary statistics for overall grade distributions, college-wise, department-wise, and course-level. Examine mean, median, standard deviation, and any potential outliers.

***Pattern Analysis***

3.1 Clustering

Utilize clustering techniques to group colleges, departments, and courses based on similar grade outcomes. Evaluate if specific clusters show distinct patterns during the pandemic.

3.2 Classification

Apply classification algorithms to identify factors influencing grade outcomes. Investigate if colleges, departments, or specific course attributes significantly contribute to predicting grade results.

3.3 Anomaly Detection

Identify anomalies in grade distributions, focusing on courses or departments that experienced unexpected changes during the pandemic. Conduct a fine-grained inspection of these anomalies.

3.4 Time Series Decomposition

Decompose the time series data to identify trends, seasonality, and residual components. Analyze if any irregularities in the residual component coincide with the pandemic period.

3.5 Decision Tree Analysis

Employ decision tree analysis to understand the hierarchy of factors influencing grade outcomes.

***Predictive Modeling***

4.1 Feature Selection

Select relevant course attributes for predictive modeling. Use domain knowledge and statistical methods to choose features that significantly impact grade outcomes.

4.2 Model Development

Build predictive models using machine learning algorithms such as regression or classification. Split the data into training and testing sets to evaluate model performance.

4.3 Model Evaluation

Assess the accuracy, precision, recall, and F1-score of the developed models. Compare models' performance across different colleges, departments, and course levels.

4.4 Interpretation

Interpret the results of the predictive models to understand the key factors influencing grade outcomes. Provide insights into how colleges, departments, and course attributes contribute to student success.

***Reporting***

5.1 Prepare a comprehensive report summarizing the methodology, findings, recommendations and potential future research. This report will include visualizations such as charts and graphs to enhance the presentation of results. Clearly communicate the findings of the study in an effort to student support services and academic planning at the University of Arizona.

5.2 Future Research

Highlight areas for future research based on gaps identified during the analysis. This could involve further investigation into specific anomalies or exploring the long-term effects of the pandemic on student success.

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

**Week 1 Nov 1st - Nov 7th**

-   Finalize proposal - group

-   Clean and explore data - group

-   Assign roles and responsibilities

**Week 2 Nov 8th - Nov 14th**

-   Exploratory data analysis

-   Continue data preprocessing.

-   Conduct exploratory data analysis (EDA) to identify initial trends and patterns.

**Week 3 Nov 15th - Nov 21st**

-   Anomaly and time series analysis 

**Week 4 Nov 22nd - Nov 28th**

-   Decision Tree Analysis

**Week 5 Nov 29th - Dec 5th**

-   (Ambitious goal :))Predictive modeling if time permits

**Week 6 Dec 6th - 12th**

-   Finalize the project and prepare for the presentation
