---
title: "Descriptive analysis of grade outcomes"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
```

DEW Counts Data Set

The DEW counts data set contains information about the breakdown of grade value outcomes during the academic years of 18-19, 19-20, and 20-21. The consists of 35697 observations with 23 variables per observation. Below is a description of the data:

| Column Name                  | Column Description                                                                         |
|------------------------------|--------------------------------------------------------------------------------------------|
| College                      | Identifies the college at the University of Arizona                                        |
| Department                   | Identifies which department within given college                                           |
| Subject.Code                 | Identifies the subject of the course                                                       |
| Catalog.Number               | Identifies the catalog number of the course                                                |
| Course.Description           | Provides the title of the course                                                           |
| Course.Level                 | Identifies if the course is lower, upper or graduate division                              |
| Total.Student.Count          | The total enrollment for the course                                                        |
| D_GRADE_COUNT                | The number of enrolled students that earned a D grade in the course                        |
| FAIL_GRADE_COUNT             | The number of enrolled students that earned a E(Fail) grade in the course                  |
| WITHDRAW_GRADE_COUNT         | The number of enrolled students that withdrew from the course                              |
| DEW_COUNT                    | The total of D, E, and W grades earned by enrolled students in the course                  |
| PASS_GRADE_COUNT             | The number of enrolled students that received a C or greater grade value                   |
| WITHDRAW_FULLMED_GRADE_COUNT | The number of enrolled students that took a full medical withdrawal                        |
| INCOMPLETE_UNGRADED_COUNT    | The number of enrolled students that did not receive grades                                |
| TERM_LD                      | The semester that the course was offered                                                   |
| ACAD_YR_SID                  | The academic year that the course was offered                                              |
| Percent.D.Grade              | The percentage of enrolled students that received a D grade in the course                  |
| Percent.E.Grade              | The percentage of enrolled students that received a E grade in the course                  |
| Percent.W.Grade              | The percentage of enrolled students that received a W grade in the course                  |
| Percent.DEW                  | The percentage of enrolled students that received a D, E, or W grade in the course         |
| Percent.Passed               | The percentage of enrolled students that received a C or greater grade value in the course |
| Per.Full..Medical.Withdrawal | The percentage of enrolled students that had a full medical withdraw in the course         |
| Per.Ungraded..Incomplete     | The percentage of enrolled students that did not receive a grade value in the course       |

```{r}
# set working directory
setwd("~/R/project-final-The-Null-Wranglers")
# read in data
dew_counts <- read.csv("data/DEW Rates.csv")
dew_counts
```

## Questions

1.  What if any patterns or relationships exist between colleges, departments, or course level and grade outcomes for enrolled students?

2.  Can a predictive model be developed that accurately forecasts grade outcomes based on course attributes?

#### **Why these questions?**

Low student success rates have a profound impact on an educational institution's academic offerings, their ability to provide financial support and scholarships, as well as their overall professional reputation. The impacts of covid-19 on student success is a concern in higher education. Thus, it is important that this vital issue be addressed with unique strategies and innovative plans that are based on research that is data driven.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

**Week 1 Nov 1st - Nov 7th**

-   Finalize proposal - group

-   Clean and explore data - group

-   Assign roles and responsibilities

**Week 2 Nov 8th - Nov 14th**

-   Exploratory data analysis

-   Continue data preprocessing.

-   Conduct exploratory data analysis (EDA) to identify initial trends and patterns.

**Week 3 Nov 15th - Nov 21st**

-   Anomaly and time series analysis 

**Week 4 Nov 22nd - Nov 28th**

-   Decision Tree Analysis

**Week 5 Nov 29th - Dec 5th**

-   (Ambitious goal :))Predictive modeling if time permits

**Week 6 Dec 6th - 12th**

-   Finalize the project and prepare for the presentation
