---
title: "Descriptive analysis of grade outcomes"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

## Dataset

DEW Counts Data Set

The DEW counts data set contains information about the breakdown of grade value outcomes during the academic years of 18-19, 19-20, and 20-21. The consists of 35697 observations with 23 variables per observation. Below is a description of the data:

| Column Name                  | Column Description                                                                         |
<<<<<<< HEAD
|-------------------|-----------------------------------------------------|
=======
<<<<<<< HEAD
|------------------------------|--------------------------------------------------------------------------------------------|
=======
|--------------------|----------------------------------------------------|
>>>>>>> 7e45ef850f1ab3d0b803aaa0fd244c7460ff35a0
>>>>>>> 44ae21815e6d6370a06a981802428fa5b712eaee
| College                      | Identifies the college at the University of Arizona                                        |
| Department                   | Identifies which department within given college                                           |
| Subject.Code                 | Identifies the subject of the course                                                       |
| Catalog.Number               | Identifies the catalog number of the course                                                |
| Course.Description           | Provides the title of the course                                                           |
| Course.Level                 | Identifies if the course is lower, upper or graduate division                              |
| Total.Student.Count          | The total enrollment for the course                                                        |
| D_GRADE_COUNT                | The number of enrolled students that earned a D grade in the course                        |
| FAIL_GRADE_COUNT             | The number of enrolled students that earned a E(Fail) grade in the course                  |
| WITHDRAW_GRADE_COUNT         | The number of enrolled students that withdrew from the course                              |
| DEW_COUNT                    | The total of D, E, and W grades earned by enrolled students in the course                  |
| PASS_GRADE_COUNT             | The number of enrolled students that received a C or greater grade value                   |
| WITHDRAW_FULLMED_GRADE_COUNT | The number of enrolled students that took a full medical withdrawal                        |
| INCOMPLETE_UNGRADED_COUNT    | The number of enrolled students that did not receive grades                                |
| TERM_LD                      | The semester that the course was offered                                                   |
| ACAD_YR_SID                  | The academic year that the course was offered                                              |
| Percent.D.Grade              | The percentage of enrolled students that received a D grade in the course                  |
| Percent.E.Grade              | The percentage of enrolled students that received a E grade in the course                  |
| Percent.W.Grade              | The percentage of enrolled students that received a W grade in the course                  |
| Percent.DEW                  | The percentage of enrolled students that received a D, E, or W grade in the course         |
| Percent.Passed               | The percentage of enrolled students that received a C or greater grade value in the course |
| Per.Full..Medical.Withdrawal | The percentage of enrolled students that had a full medical withdraw in the course         |
| Per.Ungraded..Incomplete     | The percentage of enrolled students that did not receive a grade value in the course       |

```{r}
#| label: load-dataset
#| message: false
# set working directory
setwd("~/R/project-final-The-Null-Wranglers")
# read in data
dew_counts <- read.csv("data/DEW Rates.csv")
dew_counts
```

## Questions

1.  What if any patterns or relationships exist between colleges, departments, or course level and grade outcomes for enrolled students?

2.  Can a predictive model be developed that accurately forecasts grade outcomes based on course attributes?

#### **Why these questions?**

Low student success rates have a profound impact on an educational institution's academic offerings, their ability to provide financial support and scholarships, as well as their overall professional reputation. The impacts of covid-19 on student success is a concern in higher education. Thus, it is important that this vital issue be addressed with unique strategies and innovative plans that are based on research that is data driven. By examining the grade value outcomes for the years before, during and after the pandemic, patterns and trends may be discovered that might help guide student support services. The analysis will breakdown the grade values outcomes by college, department, division level and course level using clustering and classification techniques. Anomalies will be isolated for a more fine grained inspection that includes course attributes. These attributes will be used in an effort to create a predictive model for future use.

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).

**Week 1 Nov 1st - Nov 7th**

-   Finalize proposal - group

-   Clean and explore data - group

-   Assign roles and responsibilities

**Week 2 Nov 8th - Nov 14th**

-   Exploratory data analysis

-   Continue data preprocessing.

-   Conduct exploratory data analysis (EDA) to identify initial trends and patterns.

**Week 3 Nov 15th - Nov 21st**

-   Anomaly and time series analysis 

**Week 4 Nov 22nd - Nov 28th**

-   Decision Tree Analysis

**Week 5 Nov 29th - Dec 5th**

-   (Ambitious goal :))Predictive modeling if time permits

**Week 6 Dec 6th - 12th**

-   Finalize the project and prepare for the presentation
